---
title: "Introduction to R"
subtitle: "Homework 9"
author: "Koech Andrew"
R version: "`r R.version.string`"
# date: "`r format(Sys.Date(), '%d %B, %Y')`"
date: "`r Sys.time()`"
mainfont: Calibri Light
output: pdf_document
geometry: margin=1in

header-includes:
  - \usepackage{amsfonts}
  - \usepackage{amssymb}
  - \usepackage{mathtools}
  - \usepackage{braket}

---
\newpage

## Homework task 1

When treating a condition, it is said, post-operative complications are known to occur in 25% of cases. A surgeon tests a new procedure on 10 patients and no complications occur. Is the new method better? Use the significance level $\alpha= 0.05$.

```{r,message=FALSE,warning=FALSE,include=TRUE}
#Ho:p=>0.25 vs H1:p < 0.25
binom.test(0, 10, p = 0.25, alternative = "l",0.95)
#We fail to reject the null hypothesis at alpha=0.05 and conclude the new method is not statistically better
```
<!-- EV: 4/4 -->


## Homework task 2

In 11 subjects, platelet aggregation was measured before and after smoking a cigarette. 

a) Assume that the values within the before and after smoking groups, respectively, are realizations of independent, identically normally distributed random variables. Use a suitable t-test (at significance level $\alpha= 0.05$) to test the alternative hypothesis that nicotine is expected to increase aggregation.

b) Use the Wilcoxon-test and repeat the test in a).


```{r,message=FALSE,warning=FALSE,include=TRUE}
#Ho:u2-u1=<0 vs H1:u2-u1>0
Before<-c(25,25,27,44,30,67,53,53,52,60,28)
After<-c(27,29,37,56,46,82,57,80,61,59,43)
var(Before);var(After);sd(Before);sd(After);
#Therefore we assume the variances are different; hence use a paired t-test since the subjects are the same overall
t.test(Before,After,alternative="greater",paired=TRUE,var.equal = FALSE)

## EV: If you state it like here you test whether before is "greater" than after,
# but you want to show an increase which would be the other way around:
t.test(Before,After, mu=0, alternative = "less",paired = T,conf.level = 0.95) # test: before 'less' than after
# or
t.test(After,Before, mu=0, alternative = "greater",paired = T,conf.level = 0.95) # test: after 'greater' than before (same as the other)



#We fail to reject the null hypothesis at alpha=0.05 and conclude that their exist no statistically significant difference to infer that nicotine is expected to increase  platelet aggregation
wilcox.test(Before,After,alternative="greater",paired=TRUE)
wilcox.test(After,Before,alternative="greater",paired=TRUE) ## EV: as above
#We also fail  to reject the null hypothesis at alpha=0.05 
```
<!-- EV: 2.5/4 -->

## Homework task 3

Load the data set _birthwt_ from the package MASS and read the help function for this data.

a) Using these data, use a t-test (at significance level $\alpha= 0.05$) to test the alternative hypothesis that a newborn's weight is dependent on her mother's smoking behavior.

b) Test the alternative hypothesis that the weight of a newborn is dependent on the blood pressure (ht) of its mother using a t-test (at significance level $\alpha= 0.05$) on the basis of these data.

c) Answer questions a) and b) with the Wilcoxon test.

```{r,message=FALSE,warning=FALSE,include=TRUE}
library(MASS)
data("birthwt")
attach(birthwt)
#Test used is two sided (Welch Test) since we are not interested in a specific direction so we might as well test on either directions and the groups are not of equal size
t.test(bwt~smoke,data = birthwt,alternative="t",var.equal=FALSE)
#We reject the null hypothesis at alpha=0.05 (p-value<0.05) and conclude that their exist a statistically significant dependence of a newborn's weight on her mother's smoking behavior
t.test(bwt~ht,data = birthwt,alternative="t",var.equal=FALSE)
#We fail to reject the null hypothesis at alpha=0.05 (p-value>0.05) and conclude that their exist no statistically significant dependence of a newborn's weight on her mother's blood pressure
wilcox.test(bwt~smoke,data = birthwt,alternative="t")
#As with the t-test we reject the null hypothesis at alpha=0.05 (p-value=0.006768) and conclude that their exist a statistically significant dependence of a newborn's weight on her mother's smoking behavior
wilcox.test(bwt~ht,data = birthwt,alternative="t")
#We also fail to reject the null hypothesis at alpha=0.05 (p-value = 0.1176) and conclude that their exist no statistically significant dependence of a newborn's weight on her mother's blood pressure
```
<!-- EV: 6/6 -->

## Homework task 4

The weight gain of rats is investigated under different feeding regimes:

a) Use a t-test (or Welch-Test) to test the alternative hypothesis ($\alpha= 0.05$), that feeding with high protein content leads to a higher weight gain. Assume that the observations come from normal distributions.

b) Answer this question with the Wilcoxon test.

```{r,message=FALSE,warning=FALSE,include=TRUE}
lowprotein<-c(66,112,96,78,111,121,95)
highprotein<-c(130,141,99,120,116,144,105,94,110)
sd(lowprotein);sd(highprotein)
t.test(lowprotein,highprotein,alternative="g",var.equal=FALSE,conf.level = 0.95)
#We fail to reject the null hypothesis at alpha=0.05 (p-value = 0.975) and conclude that their exist no statistically significant difference to infer that feeding with high protein content leads to a higher weight gain

## EV: Here, you have to be careful again with the direction of the test. 
# Like this you test whether the low protein values are greater than the highprotein values. Changing this leads to
t.test(highprotein,lowprotein,alternative="g",var.equal=FALSE,conf.level = 0.95)
# (Are the highprotein values actually greater than the lowprotein values?)
## So you actually get a p-value <0.05 and thus can reject the null. 

wilcox.test(lowprotein,highprotein,alternative="g",var.equal=FALSE,conf.level = 0.95) ## EV: As above.
wilcox.test(lowprotein,highprotein,alternative="l",var.equal=FALSE,conf.level = 0.95) ## EV: As above. This time p-value > 0.05.

#We fail to reject the null hypothesis at alpha=0.05 (p-value = 0.9546) and conclude that their exist no statistically significant difference to infer that feeding with high protein content leads to a higher weight gain
```

<!-- EV: 4/6 -->

## Extra Task:

For n = 50 and k = 30 (number of successes) determine the smallest probability $p_0$, for which the null hypothesis H0 : $p \ge p_0$ can still just be rejected at a significance level of 5% by the one-sided lower binomial test (that is H1 : $p < p_0$).

```{r,message=FALSE,warning=FALSE,include=TRUE}
kp0<-0#Initializing the smallest prob of rejecting
x<-seq(0,1,0.0001)#Vector of all possible prob values between 0 and 1
for (p in x) {
  decision<-binom.test(30, 50, p, alternative="l", conf.level=0.95)
  if (decision$p.value<=0.05) {
    kp0<-p
    break
  }
}
print(kp0)

binom.test(30, 50, p=0.7169, alternative="l", conf.level=0.95)#Checking validity of the result
```

<!-- EV: 4/4 -->

<!-- Total 20.5/20 -->