---
title: "Introduction to R"
subtitle: "Homework 10"
author: "Koech Andrew"
R version: "`r R.version.string`"
# date: "`r format(Sys.Date(), '%d %B, %Y')`"
date: "`r Sys.time()`"
mainfont: Calibri Light
output: pdf_document
geometry: margin=1in

header-includes:
  - \usepackage{amsfonts}
  - \usepackage{amssymb}
  - \usepackage{mathtools}
  - \usepackage{braket}

---
\newpage

## Homework task 1

In a study it will be shown that the blood glucose value can be reduced by $10mg/dl$ under a certain therapy. From preliminary studies $\sigma = 20mg/dl$ is known. The test is to be performed with a significance level 0.05 and a power of 0.90. Calculate the sample size needed to confirm this effect.

A study of 40 patients is being conducted. How large should the actual effect (blood glucose reduction) be to be detected with a significance level of 0.05 and a power of 0.90. Assume that $\sigma = 10mg/dl$.


```{r,warning=FALSE,include=TRUE,message=FALSE}
# ?power.t.test()
sn<-power.t.test(n = NULL,delta =10,sd=20,sig.level = 0.05,power = .90)
ceiling(sn$n)*2#n=172 for both groups altogether

min.delta<-power.t.test(n = 40,delta =NULL,sd=10,sig.level = 0.05,power = .90)
min.delta$delta#7.339195
```

<!-- EV: 6/6 -->

## Homework task 2

Load the dataset _iris_. Using this dataset, we want to test the null hypothesis that the means of Sepal.Length are the same for different Species, i.e. ($\mu_{versicolor} = \mu_{setosa} = \mu_{virginica}$). Assume that Sepal.Length is normally distributed. Use the pairwise t test with Bonferroni and Holm corrections with $\alpha = 0.05$ for your test. Execute the same procedure for the other metric-scaled features in the dataset.

```{r,warning=FALSE,include=TRUE,message=FALSE}
#H0:Sepal.Length are the same for versicolor,setosa,virginica Vs H1: Sepal.Length are different for atleast one species
data("iris")
attach(iris)

pairwise.t.test(Sepal.Length,Species,pool.sd = FALSE, p.adjust.method = "holm")
pairwise.t.test(Sepal.Length,Species,pool.sd = FALSE, p.adjust.method = "bonferroni")
#According to both corrections the p-values<0.05 hence we can conclude that the means of Sepal Length are significantly different for the three Species
pairwise.t.test(Sepal.Width,Species,pool.sd = FALSE, p.adjust.method = "holm")
pairwise.t.test(Sepal.Width,Species,pool.sd = FALSE, p.adjust.method = "bonferroni")
#According to both corrections the p-values<0.05 hence we can as well conclude that the means of Sepal Width are significantly different for the three Species
pairwise.t.test(Petal.Length,Species,pool.sd = FALSE, p.adjust.method = "holm")
pairwise.t.test(Petal.Length,Species,pool.sd = FALSE, p.adjust.method = "bonferroni")
#According to both corrections the p-values<0.05 hence we can as well conclude that the means Petal Length are significantly different for the three Species
pairwise.t.test(Petal.Width,Species,pool.sd = FALSE, p.adjust.method = "holm")
pairwise.t.test(Petal.Width,Species,pool.sd = FALSE, p.adjust.method = "bonferroni")
#According to both corrections the p-values<0.05 hence we can as well conclude that the means Petal Width are significantly different for the three Species
```

<!-- EV: 6/6 -->

## Homework task 3

```{r,warning=FALSE,include=TRUE,message=FALSE}
#part 1
ST<-c("NAME: Maria /COUNTRY:uruguay /EMAIL: mariaUY@gmail.com")

split.part1<-strsplit(ST,split = " /")[[1]]
split.part2<-strsplit(split.part1,":")
result<-matrix(unlist(split.part2),ncol=2,byrow=T);result

##### EV: 2/2 #####

#part 2
Position <- function(text){
  x <- letters
  pos <- match(strsplit(text,"")[[1]], x)
  return(pos)
}
Position("abcdza")

##### EV: 2/2 #####

#part 3
library(ggplot2)
data("diamonds")
attach(diamonds)

# ggplot(diamonds, aes(x=carat, y=price,col=color)) + 
#   geom_point() +  labs(x = "Carat", y = "Price")

plot(price~carat,pch=19,col=color,ylim=c(0,20000),xlim=c(0,6),
     main="Plot of Price versus Carat in Diamonds",
     xlab = "Carat", ylab = "Price")
grid(lty = "dotted",col = "lightgray")
points(price ~ carat, pch = 19, col = color)
legend("topright", legend = levels(color), pch = 19,
       col = 1:length(levels(color)),title = "Color")


mod<-lm(price~carat,data = diamonds)
abline(mod,col="red",lwd=2)

##### EV: 2/2 #####
```

<!-- Total 18/20 -->
