---
title: "Introduction to R"
subtitle: "Homework 11"
author: "Koech Andrew"
R version: "`r R.version.string`"
# date: "`r format(Sys.Date(), '%d %B, %Y')`"
date: "`r Sys.time()`"
mainfont: Calibri Light
output: pdf_document
geometry: margin=1in

header-includes:
  - \usepackage{amsfonts}
  - \usepackage{amssymb}
  - \usepackage{mathtools}
  - \usepackage{braket}

---
\newpage

## Homework task 1

Test whether there is a relationship between smoking and gender using the Fisher's exact test and the $\chi^2$-test.

```{r}
observed.freq<- matrix(c(14,26,1,12),2,byrow = F)
colnames(observed.freq) <- c("non-smokers","smokers")
rownames(observed.freq) <- c("men","women");observed.freq

chisq.test(observed.freq)#Since the p-value > 0.05 we conclude that their exists no statistically significant relationship between smoking and gender
fisher.test(observed.freq)#Since the p-value > 0.05 we conclude that their exists no statistically significant relationship between smoking and gender
DescTools::GTest(observed.freq)#Since the p-value < 0.05 we conclude that their exists a statistically significant relationship between smoking and gender
```

<!-- EV: 5/5 -->


## Homework task 2

In the _survey_ dataset from the MASS package, consider the variables Pulse (pulse rate, beats per minute) and Smoke (smoking behavior, different categories). Is there a statistically significant relationship between pulse rate and smoking behavior? Use the cor.test function for a test based on Kendall's tau. Consider the variable Smoke as numerical, i.e. assign numerical values 1 to 4 to the expressions Never, Occas, Regul and Heavy in a meaningful order. $(Never< Occas < Regul < Heavy)$ Use $\alpha= 0.05$  for your tests.

```{r}
library(MASS)
data("survey")
attach(survey)
Smoke<-as.numeric(factor(Smoke,levels = c("Never","Occas","Regul","Heavy")))
cor.test(Pulse,Smoke,method = "kendall")
#There exists no statistically significant relationship between pulse rate and smoking behaviour at alpha=0.05 since the p-value = 0.8301 > 0.05
```

<!-- EV: 5/5 -->

## Homework task 3

```{r}
'%dsum%'<-function(x,y){
2*(x+y)  
}

3%dsum%5
2%dsum%7

## EV: colon operator missing.
```
<!-- EV: 2/3 -->

## Homework task 4

```{r}
set.seed(528164)
library(tidyverse)
library(fmsb)
library(gridExtra)
field1<-runif(3,0,20) 
field2<-runif(3,0,20) 
field3<-runif(3,0,20) 
field4<-runif(3,0,20) 
field5<-runif(3,0,20)
max.value<-rep(20,5)
min.value<-rep(0,5)
skills<-data.frame(field1,field2,field3,field4,field5)
df<-rbind(max.value,min.value,skills)
rownames(df)<-c("max.value","min.value","namea","nameb","namec")

radarchart(df)

op <- par(mar = c(1,2, 4, 12))
layout(matrix(c(1, 2), nrow = 1), widths = c(3, 1))
radarchart(df, plwd=2, plt=2, cglcol="grey", cglty=2, axistype =0,
           pfcol ="#99999980",pcol =c("#00AFBB", "#E7B800", "#FC4E07"))

legend(x = "bottom", legend = rownames(df[-c(1,2),]), horiz = TRUE,bty = "n",
       pch = 20 , col = c("#00AFBB", "#E7B800","#FC4E07"))

df=df[-c(1,2),]
df <- df %>%
  rownames_to_column(var = "student_names") %>%
  pivot_longer(-student_names, names_to = "field", values_to = "Value")

ggplot(df, aes(factor(field), Value, fill = student_names)) +
  geom_bar(stat="identity", position = "stack") +
  scale_fill_brewer(palette = "Set1")+
  labs(x="field",y="Skill level")

par(op)

```

<!-- EV: 7/7 -->

<!-- Total 19/20 -->
